---
title: "Biodiversity Analysis"
author: "Maya Craig"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.path = "../figures/04_Biodiversity")
```

# Setting the Environment 

## Set the seed 
```{r}
set.seed(76544)
```

## Load Libraries 
```{r load-packages}
pacman::p_load(tidyverse, devtools, patchwork, iNEXT, phyloseq, install = FALSE)
```


## Load in Data 
```{r}
load("data/02_PreProcessing/raw_preprocessed_physeq.RData")
raw_preprocessed_physeq

# Intution Check 
min(sample_sums(raw_preprocessed_physeq))

# Make metadata dataframe
metadata_df <- 
  raw_preprocessed_physeq %>%
  sample_data() %>%
  data.frame()

# setting colors for the stations
station_colors <- c(
  "Shipping Channel" = "dodgerblue4",
  "Aransas Bay" = "dodgerblue2",
  "Copano West" = "#D9CC3C",
  "Copano East" = "#A0E0BA",
  "Mesquite Bay" = "#00ADA7")
```

# Goals 
1. Calculate the Hill Diversity of the samples 
2. Evaluate the rarefaction curves
3. Evaluate the Diversity values
4. Makes notes of the specific samples and their seq depth

# Diversity Calculations with iNEXT

```{r}
#prepare input 
iNEXT_input_df <-
raw_preprocessed_physeq %>%
  otu_table() %>%
  data.frame()

# Quick Check
dim(iNEXT_input_df)

# Run iNEXT: Calculate the Hill numbers
# Note that: Species in ROWS, Samples in Columns

# Remember to set the seed!
#iNEXT_data <- iNEXT(iNEXT_input_df, q = c(0,1,2), datatype = "abundance")

# Save the file 
#save(iNEXT_data, file = "data/04_Biodiversity/iNEXT_data.RData")
```
q = 0; is the richness (total number of species)
q = 1; the Shannon -- emphasizes more common taxa 
q = 2; Simpson -- emphasized more dominant taxa


# Evaluate the Diversity
```{r load-div}
load("data/04_Biodiversity/iNEXT_data.RData")
str(iNEXT_data)
typeof(iNEXT_data)
```
- the iNext package does interpolation (calculate were the max amout of sequences are in a sample) and extrapolation to 2x the library size 
- if the observed value is less than the extrapolated value, then we should consider rarefication because there is not quite enough data 

# Plot Diversity
```{r plot-rarefaction}
# Prepare Colors 
color_df <- iNEXT_input_df %>%
  colnames() %>%
  data.frame()

#Check
head(color_df)
# Rename the column 
colnames(color_df)[1] <- "names"
# Check
head(color_df)

# Make a helper dataframe for plotting with colors 
color_df %>%
  mutate(names = gsub(names, pattern = "[.]", replace = "-"), 
         names = gsub(names, pattern = "X", replace = "")) %>% #fix names for merging 
  left_join() %>%
  head()
```



